---
title: "Superefficient teamwork in weaver ants"
author: "Code by: Daniele Carlesso and Madelyne Stewardson"
date: "2025-07-01"
output: 
  html_document: 
    fig_width: 10
    fig_height: 10
    toc: true
---

# . Set working directory. Load packages. Set theme for figures

```{r,error=FALSE}

setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) ## Data must be in same folder

if (!require(librarian)) {
  install.packages("librarian")
  library(librarian)
}

shelf(tidyverse, ggplot2, ggpmisc, viridis, ggpubr, lme4, lmerTest, DHARMa, sjPlot, glmmTMB, effects, ggeffects, AICcmodavg, modelsummary, emmeans, data.table, MuMIn, Durga, car, truncnorm, random, performance, png, grid, loo, bayesplot, brms, png, grid, gtools, ggnewscale)

set_theme(base = theme_bw(),
          axis.title.size = 1.5,
          axis.textsize = 1.0, axis.title.color = 'black', axis.textcolor = 'black',
          legend.title.size = 1.5)

```

# . Load datasets

df = dataset on chain size and force
allchainposdf = ant posture dataset
thoraxDF = thorax length dataset

```{r}

df <- read.csv("ChainData.csv") 
allchainposdf <- read.csv("ChainPos.csv") 
thoraxDF <- read.csv("ThoraxLengths.csv")

```

# . Summarise dataset by arrangement and calculate chain length ratio

Arrangement changes every time an ant joins or leaves the pulling effort. Note that 'Chain length ratio' is named 'propmax' throughout the code.

```{r}

meandf <- df %>%
  group_by(replicate, newarr) %>%
  mutate(n_row = n()) %>% ## Duration of arrangement in seconds
  mutate(
    state = ifelse(ptime < 0, 'grow', 'decay'),
    meanforceN = mean(forceN),
    meanforceperantN = mean(forceperantN)) %>%
  slice_head() %>%
  mutate(
    prop1 = chain1*1/total.ants,
    prop2 = chain2*2/total.ants,
    prop3 = chain3*3/total.ants,
    prop4 = chain4*4/total.ants,
    prop5 = chain5*5/total.ants,
  ) %>%
  mutate(
    propmax = as.factor(which.max(c(prop1, prop2, prop3, prop4, prop5))), ## chain length ratio
    state = factor(state, levels = c("grow","decay"))
  )

```

### . Summary of experimental observations

```{r}

length(unique(meandf$replicate)) ## 15 reps
length(unique(meandf$arrangement)) ## 108 arrangements
length(unique(meandf$colony)) ## 5 colonies
n <- meandf %>% group_by(replicate) %>% summarise(n = length(unique(newarr))); sum(n$n) ## 364 new arrivals
nrow(meandf) ## 364 observations in total

```

### . Average force produced by ants

```{r}

avg_forceperant <- mean(meandf$meanforceperantN) ## Average individual force output throughout all replicates

sem_forceperant <- sd(meandf$meanforceperantN)/sqrt(length(meandf$meanforceperantN)) ## Standard deviation

max_forceperant <- max(meandf$meanforceperantN) ## Maximum individual force output recorded

mean(meandf[meandf$total.ants == 1,]$meanforceperantN) ## Mean individual force output when team size == 1

mean(meandf[meandf$total.ants == 15,]$meanforceperantN) ## Mean individual force output when team size == 15

```


# . Data analysis and visualization

## . GLMM on team force output

```{r}

# Fit model
fit1 <- glmmTMB(sqrt(meanforceN) ~ total.ants * state + propmax + (1|colony/replicate), data = meandf) 

# Diagnostics
check_model(fit1)

# Results
summary(fit1)

Anova(fit1, "III")

# Forest plot
plot_model(fit1)

## Pairwise comparison between chain length ratio 
emms <- emmeans(fit1, ~ propmax)
pairs(emms)

### PLOTTING ###

## Predict based on total number of ants and state
dat <- ggpredict(fit1, terms = c("total.ants [all]", "state [all]")) %>%
  rename(state = group)

## Visualise results 
FigX1 <- ggplot() +
  geom_point(data = meandf, aes(total.ants, meanforceN, color = state, fill = state), size = 3) +
  geom_line(data = dat, aes(x, predicted, color = state)) +
  geom_ribbon(data = dat, aes(x = x, ymin = conf.low, ymax = conf.high, fill = state), alpha = .4) +
  labs(
    x = "Total nº of ants",
    y = "Total Force (mN)",
    fill = "Phase",
    colour = "Phase") +
  scale_color_manual(values = c("#00BFC4", "#F8766D"), labels = c("Growth", "Decay")) +
  scale_fill_manual(values = c("#00BFC4", "#F8766D"), labels = c("Growth", "Decay")) +
  theme_pubr(legend = c(0.1, 0.9),
             base_size = 13); FigX1


## Effect without transformation
dat1 <- ggeffect(fit1, terms = c("total.ants [all]")) %>%
  rename(state = group) 

ggplot(meandf) +
  geom_point(aes(total.ants, sqrt(meanforceN), color = state)) +
  geom_line(data = dat1, aes(x, predicted)) +
  geom_ribbon(data = dat1, aes(x = x, ymin = conf.low, ymax = conf.high), alpha = .2) +
  labs(
    x = "Total nº of ants",
    y = "Square-root Mean Force (mN)"
  )

## Durga plot
d <- Durga::DurgaDiff(meandf, "meanforceN", "propmax")
DurgaPlot(d, contrasts = c("2 - 1", "3 - 2", "4 - 3"), xlab = "PropMax",left.ylab = "Mean force (mN)") 

```

## . GLMM on individual force output

```{r}

# Fit model
fit2 <- glmmTMB(sqrt(meanforceperantN) ~ total.ants * state + propmax + (1|colony/replicate), data = meandf) 

# Diagnostics
check_model(fit2)

# Results
summary(fit2)
Anova(fit2, "III")

# Forest plot
plot_model(fit2)

# Pairwise comparison between chain length ratio 
emms.ifo <- emmeans(fit2, ~ propmax)
pairs(emms.ifo)


### PLOTTING ###
dat2 <- ggpredict(fit2, terms = c("total.ants [all]", "state [all]")) %>%
  rename(state = group)

## Effect of total number of ants and state
FigX2 <- ggplot() +
  geom_jitter(data = meandf, aes(total.ants, meanforceperantN, color = state), height = 0, width = 0) +
  scale_x_continuous(breaks = c(1, 5, 10, 15)) +
  geom_line(data = dat2, aes(x, predicted, color = state)) +
  geom_ribbon(data = dat2, aes(x = x, ymin = conf.low, ymax = conf.high, fill = state), alpha = .2) +
  labs(
    x = "Total nº of ants",
    y = "Mean Force per ant (mN)",
    colour = "Phase",
    fill = "Phase"
  ) +
  scale_color_manual(values = c("#00BFC4", "#F8766D"), labels = c("Growth", "Decay")) +
  scale_fill_manual(values = c("#00BFC4", "#F8766D"), labels = c("Growth", "Decay")) +
  theme_pubr(legend = c(0.1, 0.9), base_size = 13) ; FigX2

# Durga plot
d1 <- Durga::DurgaDiff(meandf, "meanforceperantN", "propmax")
p <- Durga::DurgaPlot(d1, contrasts = c("1 - 2", "2 - 1", "3 - 1", "4 - 1"), xlab = "", left.ylab = "Mean Force per ant (mN)", error.bars.type = "CI")

```

## . Joining and leaving dynamics 

Here we plot the joining and leaving dynamics for each replicate. Time is standardized so that each replicate starts at -1 and finishes at 1, where -1 to 0 indicates the growth phase, and 0 to 1 indicates the decay phase.

```{r}

df1 <- df %>%
  group_by(replicate) %>%
  arrange(replicate, ptime) %>%
  mutate(
    changeNants = total.ants - lag(total.ants, 1) ## Find instances where number of ants changes
  )

df1$joinLeave <- ifelse(  ## Label each event as joining or leaving
  df1$changeNants == 1, "Join",
  ifelse(df1$changeNants == -1, "Leave", NA) 
)

df1$timeJoinLeave <- ifelse(
  !is.na(df1$joinLeave), df1$ptime, NA
)

JL_plot <- ggplot(df1, aes(ptime, forceN)) +
  geom_path() +
  geom_vline(data = df1[!is.na(df1$joinLeave),], aes(xintercept = timeJoinLeave, color = joinLeave), linetype = 2, na.rm = T) +
  facet_wrap(~replicate)

plotly::ggplotly(JL_plot) ## Interactive plot

```

## . Force change after joining or leaving event

Force change is evaluated over the 5 seconds following a joining or leaving event

```{r}

timerange = 5 ## time of interest before/after each joining / leaving event 

df1 <- df1 %>%
  group_by(replicate, newarr) %>%
  mutate(
    duration = n()
  )

JL_events <- df1[!is.na(df1$joinLeave), ]

JL_forcechange = list()
for (row in 1:nrow(JL_events)) {

  event <- JL_events[row, ] 
  
  if (any(!is.na(df1[df1$replicate == event$replicate & 
                     between(df1$time, event$time - timerange, event$time - 1),]$joinLeave)) |
      any(!is.na(df1[df1$replicate == event$replicate & 
                     between(df1$time, event$time + 1, event$time + timerange),]$joinLeave)) |
      length(df1[df1$replicate == event$replicate & 
                     between(df1$time, event$time - timerange, event$time - 1),]$joinLeave) != timerange |
      length(df1[df1$replicate == event$replicate & 
                     between(df1$time, event$time + 1, event$time + timerange),]$joinLeave) != timerange) {
    next
    }
  
  force_pre <- mean(df1[df1$replicate == event$replicate & 
                     between(df1$time, event$time - timerange, event$time),]$forceN) 
  
  force_post <- (df1[df1$replicate == event$replicate & 
                     between(df1$time, event$time + 1, event$time + timerange),]$forceN)
  
  force_post <- force_post - force_pre[1]
  force_pre <- force_pre - force_pre[1]
  
  force <- c(force_pre, force_post)
  
  ants = event$total.ants
  ants = ifelse(between(ants, 1, 5), "1 - 5", ifelse(between(ants, 5, 10), "6 - 10", "10 +"))
  
  JL_forcechange[[length(JL_forcechange) + 1]] <- data.frame(
    replicate = rep(event$replicate[1], times = length(force_pre)),
    event = rep(event$joinLeave, times = length(force_pre)),
    ID = row,
    ants = ants,
    phase = rep(ifelse(event$ptime < 0, "Growth", "Decay"), times = length(force_pre)),
    force = force,
    time = seq(force) - 1
  )
}

JL_forcechange <- rbindlist(JL_forcechange)

JL_forcechange$phase <- factor(JL_forcechange$phase, levels = c("Growth", "Decay"))
JL_forcechange$ants <- factor(JL_forcechange$ants, levels = c("1 - 5", "6 - 10", "10 +"))

JL_forcechange_summ <- JL_forcechange %>%
  group_by(phase, event, time) %>%
  summarise(
    event = event[1],
    phase = phase[1],
    force_min = min(force),
    force_max = max(force),
    force_mean = mean(force),
    n = n(),
    SEM = sd(force)/sqrt(n()),
    sem_min = force_mean - SEM, 
    sem_max = force_mean + SEM,
    s = sd(force),
    margin = quantile(x = force, probs = 0.975)*sd(force)/sqrt(n()),
    ci_min = force_mean - margin, 
    ci_max = force_mean + margin
  )

ForceTeam_JoinLeave <- ggplot(JL_forcechange_summ, aes(time, force_mean, fill = event)) + 
  geom_ribbon(aes(ymin = sem_min, ymax = sem_max), alpha = .4) + 
  geom_line(aes(color = event)) +
  geom_point(aes(color = event)) +
  scale_fill_viridis_d(option = "H", 
                       begin = 0.1, 
                       aesthetics = c('color', 'fill')) +
  labs(
    fill = "Event",
    color = "Event",
    x = "Time from event (s)",
    y = "Team force (mN)"
  ) +
  theme(
    legend.position = c(0.9, 0.85),
    legend.title = element_text(face = "plain")
  ) +
  facet_wrap(~phase); ForceTeam_JoinLeave

```

## . Figure 2 - Main manuscript

```{r}

## Note: Durga plots do not integrate well with ggplot and ggarrange. I save it as PNG and re-upload it to combine it with other plots.

# Create temp file
tmp_file <- tempfile(fileext = ".png")

# Open a PNG device and draw the Durga plot
png(filename = tmp_file, width = 1800, height = 1300, res = 300)
Durga::DurgaPlot(d1, contrasts = c("1 - 2" = "1 - 2", "2 - 1" = "2 - 1", "3 - 1" = "3 - 1", "4 - 1" = "4 - 1"), xlab = "", left.ylab = "Mean Force per ant (mN)", error.bars.type = "CI")
dev.off()

# Read the image and convert to grob
img <- readPNG(tmp_file)
p1 <- rasterGrob(img, interpolate = TRUE, 
                 width = unit(1, "npc"), height = unit(1.1, "npc"))

pplots <- list(FigX1, p1, FigX2, ForceTeam_JoinLeave)

# Combine plots
ggarrange(plotlist = pplots, nrow = 2, ncol = 2, labels = c("A", "C", "B", "D"))

```

## . Bayesian model for force change following join/leave event

```{r}

# Convert variables into factors
JL_forcechange$timeF <- as.factor(JL_forcechange$time)
JL_forcechange$ID <- as.factor(JL_forcechange$ID)
JL_forcechange$event <- as.factor(JL_forcechange$event)

JL_df <- JL_forcechange[JL_forcechange$time == 5, ]

# Data show extreme leptokurtosis and unimodal peak
ggplot(JL_df, aes(force)) + 
  geom_histogram() + 
  xlab("X")

# BRMS model
JL_df$force_z <- (JL_df$force - mean(JL_df$force))/sd(JL_df$force) # Centering and scaling

# Set robust prior
prior <- c(
  prior(student_t(3, 0, 10), class = "b"),  # fixed effects
  prior(gamma(2, 0.5), class = "nu")        # favors moderate tail thickness
)

# Model fit
model_brm <- brm(
  force_z ~ phase * event + (1 | replicate),
  data = JL_df,
  family = student(),
  prior = prior,
  iter = 4000, warmup = 1000, chains = 4, cores = 4,
  control = list(adapt_delta = 0.99)
)

# Model diagnostics
pp <- pp_check(model_brm, type = "dens_overlay")
pp + xlim(-20, 20)

pp_check(model_brm, type = "ecdf_overlay")         # Tails or peaks
pp_check(model_brm, type = "intervals")            # Predictive intervals per point
pp_check(model_brm, type = "stat", stat = "mad")   # Median absolute deviation

posterior_summary(model_brm, variable = "nu") # Estimated degrees of freedom
mcmc_areas(as_draws_df(model_brm), pars = "nu") # Nu estimation looks robust for heavy tails

loo(model_brm)   # Pareto k diagnostics

### Results ###  
summary(model_brm)

## Hypothesis testing 

# Do joining events cause more force increase in the growth phase than in the decay phase?
hypothesis(model_brm, "phaseDecay < 0")

# Do leaving events cause more force decrease during decay phase than growth phase?
hypothesis(model_brm, "phaseDecay + phaseDecay:eventLeave < 0")

# Does force increase for joining events during the decay phase?
hypothesis(model_brm, "Intercept + phaseDecay > 0")

# Do leave events differ from 0 in the growth phase?
hypothesis(model_brm, "Intercept + eventLeave < 0")

# Do leave events differ from 0 in the decay phase?
hypothesis(model_brm, "Intercept + phaseDecay + eventLeave + phaseDecay:eventLeave < 0")

```

## . Ant posture analysis

Visualize hindleg stretch based on ant position within the chain

```{r}

allchainposdf$ant <- factor(allchainposdf$ant, levels = c('oneone','twoone','twotwo','threeone','threetwo','threethree'))

## Rename for ease of visualization
allchainposdf$nameplot <- ifelse(allchainposdf$ant == 'oneone', "Singletons", NA)
allchainposdf$nameplot <- ifelse(allchainposdf$ant == 'twoone', "2-ant-chain (front)", allchainposdf$nameplot)
allchainposdf$nameplot <- ifelse(allchainposdf$ant == 'twotwo', "2-ant-chain (back)", allchainposdf$nameplot)
allchainposdf$nameplot <- ifelse(allchainposdf$ant == 'threeone', "3-ant-chain (front)", allchainposdf$nameplot)
allchainposdf$nameplot <- ifelse(allchainposdf$ant == 'threetwo', "3-ant-chain (middle)", allchainposdf$nameplot)
allchainposdf$nameplot <- ifelse(allchainposdf$ant == 'threethree', "3-ant-chain (back)", allchainposdf$nameplot)

heat_df <- allchainposdf %>%
  group_by(chainsize, nameplot) %>%
  summarise(
    sd_stretch = sd(meanhindstretch, na.rm = TRUE), ## SD of hindleg stretch
    meanhindstretch = mean(meanhindstretch, na.rm = TRUE)) %>%  ## mean hindleg stretch
  ungroup()

## Rearrange levels for plotting
heat_df$nameplot <- factor(heat_df$nameplot, levels = c(
  "Singletons",
  "2-ant-chain (front)", "2-ant-chain (back)",
  "3-ant-chain (front)", "3-ant-chain (middle)", "3-ant-chain (back)"
))

heat_df <- heat_df[order(heat_df$chainsize, heat_df$nameplot), ] 

# Add x/y positions to draw each ant in a spatial layout
heat_df$y <- as.numeric(as.factor(heat_df$chainsize))
heat_df$x <- ave(rep(1, nrow(heat_df)), heat_df$y, FUN = seq_along)

hindstretch_plot <- ggplot(heat_df, aes(x = x, y = -y)) +  # Flip y for top-down chain order
  geom_tile(aes(fill = meanhindstretch), width = 0.9, height = 0.9, alpha = 0.7) +
  geom_text(aes(label = paste0(round(meanhindstretch, 2), "\n± ", round(sd_stretch, 2))),
            size = 3.5, color = "black") +
  scale_fill_viridis(option = "B", name = "Extension (mm)", begin = 0.1) +
  scale_y_continuous(breaks = -unique(heat_df$y),
                     labels = paste0(unique(heat_df$chainsize), "-ant chain")) +
  theme(
    legend.position = c(0.7, 0.90),
    legend.title.position = "top",
    legend.direction = "horizontal",
    legend.key.width = unit(1.5, "cm"),      # widen color bar
    legend.key.height = unit(0.4, "cm"),     # adjust height
    legend.text = element_text(margin = margin(t = 5, b = 2)),  # space above/below labels
    legend.spacing.x = unit(0.5, "cm"),      # space between ticks
    legend.box.margin = margin(t = 5, r = 10, b = 5, l = 10),  # space around legend box
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_blank(),
    axis.title = element_blank(),
    text = element_text(size = 15)
  ) +
  scale_x_continuous(
    breaks = sort(unique(heat_df$x)),
    labels = c("Front", "Middle", "Rear")
  ) 

hindstretch_plot

```

## . Figure 3 - Main manuscript

```{r}

img <- readPNG("Images/ants.png")

g_img <- rasterGrob(img, interpolate = TRUE)

image_plot <- ggplot() + 
  annotation_custom(g_img, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  theme_void(); image_plot

ggarrange(image_plot, hindstretch_plot, ncol = 2, widths = c(0.8, 1), labels = c("A", "B"))

```

## . Force of chains v. solo+chains

```{r}

SoloVSoloChains_df <- meandf %>%
  group_by(replicate) %>%
  arrange(replicate, ptime) %>%
  mutate(
    phase = ifelse(ptime < 0, "Growth", "Decay")
  )

## Labelling
SoloVSoloChains_df$solos <- ifelse(  
  SoloVSoloChains_df$total.chains == 1, "Solos", 
  ifelse(SoloVSoloChains_df$chain1 > 0 & SoloVSoloChains_df$total.chains > SoloVSoloChains_df$chain1, "Solo + Chain", "Only chains"))

SoloVSoloChains_df$phase <- factor(SoloVSoloChains_df$phase, levels = c('Growth', 'Decay'))

SoloVSoloChains_df$solos <- factor(SoloVSoloChains_df$solos, levels = c("Solos", "Solo + Chain", "Only chains"))

SoloVSoloChains_df <- SoloVSoloChains_df[SoloVSoloChains_df$solos != "Solos",]

## Statistics

# Fit model
chainComb_mod <- glmmTMB(forceperantN ~ solos + total.ants + (1|replicate), SoloVSoloChains_df)

# Diagnostics
check_model(chainComb_mod) # Using performance

res <- simulateResiduals(chainComb_mod) # Using DHARMa
testResiduals(res)

# Results
summary(chainComb_mod)

## PLOTTING ## 
plot_model(chainComb_mod, "est") # Forest plot
 
SoloVSoloChains_plt <- ggplot(SoloVSoloChains_df[SoloVSoloChains_df$solos != "Solos",], aes(total.ants, forceperantN, color = solos)) + 
  geom_point(size = 3, alpha = 1) +
  facet_wrap(~phase) +
  labs(
    x = "Total number of ants",
    y = "Force per ant (mN)",
    color = "Arrangement"
  ) + 
  theme(
    text = element_text(size = 13),
    legend.position = "top"
  )

SoloVSoloChains_plt

```

## . Number of ants over time, with Joining and Leaving events

```{r}

## Use joining/leaving dataset
df2 <- df1 %>% 
  group_by(replicate) %>%
  arrange(replicate, ptime) 

df2$solos <- ifelse(  
  df2$total.ants == 1, "1", ifelse(between(df1$total.ants, 2, 5), "2 - 5", 
                                       ifelse(between(df2$total.ants, 6, 10), "6 - 10", 
                                              ifelse(between(df2$total.ants, 11, 13), "11 - 13", "13+"))))

## Order factor variable for plotting
df2$solos <- factor(df2$solos, levels = c("1", "2 - 5", "6 - 10", "11 - 13", "13+"))

# Preserve original names
unique_reps <- gtools::mixedsort(unique(df2$replicate)) 

df2 <- df2 %>%
  mutate(replicate = paste0("Replicate ", match(replicate, unique_reps)))

df2$replicate <- factor(df2$replicate, levels = unique(df2$replicate))

y_line <- min(df2$forceN, na.rm = TRUE) - 8
df_line <- data.frame(x = -1, xend = 1, y = y_line, yend = y_line)

# Plot
JL_time_plot <- ggplot() +
  geom_path(data = df2, aes(ptime, forceN, color = solos, group = replicate), linewidth = 0.8) +
  scale_color_brewer(palette = "Set1", name = "Nº of \nants") +
  
  new_scale_color() +
  
  geom_point(
    data = df2[!is.na(df2$joinLeave),],
    aes(x = ptime, y = -8, color = joinLeave),
    shape = "triangle",
    alpha = 0.7,
    size = 1.8
  ) +
  
  scale_color_manual(values = c("Join" = "black", "Leave" = "red"), name = "Event") +

  labs(x = "Normalised time", 
       y = "Team force (mN)", 
       color = "Event",
       shape = "Event") +
  
  ylim(-8, 135) +
  theme(text = element_text(size = 15)) +
  facet_wrap(~replicate)

JL_time_plot

```

## . Force change between start and end of each arrangement

Density plot shows that, for each arrangement, force per ant increases during growth and decreases during decay. Only 

```{r}

df_forcechange_arrangement <- df %>%
  group_by(replicate, newarr) %>%
  mutate(
    duration = n(),
    timeFirst = mean(head(forceperantN), 3), ## Average force in the first 3 seconds of arrangement
    timeLast = mean(tail(forceperantN, 3)), ## Average force in the last 3 seconds of arrangement
    forcediff = timeLast - timeFirst,
    phase = ifelse(ptime < 0, "growth", "decay"),
    phase = factor(phase, levels = c("growth", "decay"))
  ) %>%
  slice_head()

force_diff_plot <- ggplot(df_forcechange_arrangement[df_forcechange_arrangement$duration > 10, ], aes(x = forcediff, color = phase)) +
  geom_density(linewidth = 1, key_glyph = draw_key_path) +
  labs(
    x = "Difference in force per ant between start and end of each arrangement (mN)",
    color = "Phase"
  ) +
  geom_vline(aes(xintercept = 0), linetype = 2, color = "black", linewidth = 0.5) +
  scale_color_manual(values = c("#00BFC4", "#F8766D"), labels = c("Growth", "Decay")) +
  theme_pubr(legend = c(0.1, 0.9),
             base_size = 13) +
  theme(strip.text = element_blank())

force_diff_plot

```

## Thorax length v. force production

```{r}

# Avoid name conflict
meandf <- meandf %>% select(-colony)

# Rename colonies
meandf$colony_short <- ifelse(startsWith(meandf$replicate, "Col"), substr(meandf$replicate, 1, 4), substr(meandf$replicate, 1, 5))

thoraxDF$colony_short <- str_remove(unique(thoraxDF$colony), "Thorax")

# Select relevant columns
thoraxDF <- thoraxDF[, c('colony_short', 'Length')]

thoraxDF <- thoraxDF[thoraxDF$colony_short %in% unique(meandf$colony_short), ]

# Thorax length is similar across colonies
ggplot(thoraxDF, aes(colony_short, Length, fill = colony_short)) +
  geom_boxplot()

# Merge datasets
thoraxDF <- thoraxDF %>%
  group_by(colony_short) %>%
  summarise(thorax_length = mean(Length))

meandf_wThorax <- merge(meandf, thoraxDF, by = "colony_short")

map = setNames(c("Colony 1", "Colony 2", "Colony 3", "Colony 4", "Colony 5"), c(unique(meandf_wThorax$colony_short)))

meandf_wThorax <- meandf_wThorax %>%
  mutate(colony_short = map[colony_short])


## PLOTTING ##
ggplot(meandf_wThorax, aes(thorax_length, meanforceperantN)) +
  geom_point(aes(color = colony_short)) +
  geom_smooth(method = "lm") +
  labs(
    color = "Colony", 
    y = "Mean force per ant (mN)",
    x = "Thorax length (cm)"
  )

ggplot(meandf_wThorax, aes(thorax_length, chainsize)) +
  geom_point(aes(color = colony_short)) +
  geom_smooth(method = "lm") +
  labs(
    color = "Colony", 
    y = "Chain size (ants)",
    x = "Thorax length (cm)"
  )


## STATISTICS ##
mod_thorax <- glmmTMB(meanforceperantN ~ thorax_length, meandf_wThorax)

# Diagnostics
res <- simulateResiduals(mod_thorax, n = 1000)
testResiduals(res)

# Results
summary(mod_thorax)

```

